COVID Shutdown Regulations Negatively Impact Employment and Sales of Restaurants in Ontario

### 2021-02-26 outputs/paper/paper.Rmd
Commit for Sampling Method

header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(bookdown)
library(janitor)
library(stringr)
library(ggrepel)
library(here)
library(kableExtra)
```

@@

&nbsp; &nbsp; Our data explores shutdown effects on restaurant businesses in Ontario. We analyzed it using R [@citeR], and packages `tidyverse` [@cite_tidyverse], `stringr`[@cite_stringr],  `here` [@cite_here], `ggpubr`[@cite_ggpubr], `janitor`[@cite_janitor] . We used R packages `cansim` [@cite_cansim], `httr`[@cite_httr], `xml2`[@cite_xml2] to decide our intervention and sampling method, and packages `bookdown` [@cite_bookdown], `kableExtra`[@cite_kableextra], `ggrepel`[@cite_ggrepel] to format the document. We referenced `Impact Evaluation in Practice`[@cite_impact] and `Sampling Theory and Practice`[@cite_sampletheory] to design our experiment.

@@

&nbsp; &nbsp; Starting from October, 2020, the Ontario government reopened Peterborough while keeping shutdown restrictions on the rest of Ontario regions. After a three month period in January 2021, we at Petit Poll started to conduct survey on restaurants in Peterborough and Bratford about their business performance from October to December. Survey results are released in February 2021.

@@

A presurvey with five questions was conducted before the decision on intervention was made. The purpose was to make sure our assumptions about the resturant similarities between Peterborough and Brantford are based on concert evidence. We asked selected respondents about the restaurants' location, employment, sales, service types, average price per meal per person to assess whether the resturuants in the two cities are of considerable similarities.

**We considered all Ontario restaurants as its target population. The frame population was decided by randomly selecting 100 restaurants from Brantford and 100 from Peterborough. To ensure the external validity of the experiment, we used the following steps to find these restaurants: **

1. Scrape the postal codes of all restaurants located within the circle of the Brantford City Centre to a 40 km radius on Yelp. And then use the same method for Peterborough.

@@

3. Compare the two lists and eliminate the restaurants with postal codes outside of Peterborough CMA and the Brantford CMA.

4. Randomly select 100 restaurants from Brantford and 100 restaurants from Peterborough to form the frame population.

```{r presamplingtable, fig.cap='Sampling Results Summary', echo=FALSE }
tibble(
  'City' = list('Brantford', 'Peterborough'), 
  'Survey Administered' = list(100, 100),
  'Total Sample Collected' = list(50, 55),
  'Telephone Survey' = list(43, 47),
  'Online Survey' = list(7, 8)
) %>%
  kbl(caption = 'Sampling Results Summary',
      align = "cccc") %>%
  kable_styling(latex_options = "hold_position")
```

We used telephone interviews and online questionnaire methods to administer the survey. The telephone interview served as the primary channel and an online survey option was provided when respondents could not or prefer not to be interviewed over the phone. As shown in the above table (Table \@ref(tab:presamplingtable)), our sampled population included 50 valid responses in Brantford and 55 collected from Peterborough, among which, 7 were collected via the online survey in Brantford and 8 were collected via the online survey in Peterborough. Surveys that were not answered or have incomplete answers were treated as nonresponses and were not included in the dataset for further analysis. The surveys were administered in the following steps:

1. The phone number of the restaurants was scraped from Yelp while we determined the frame population.

@@

3. When the call was connected, the survey staff would provide a brief introduction of the survey purpose and indicate that the survey was sponsored by the Ontario government. See Appendix A for details.

@@

6. To address privacy concerns and to increase the response rate, the respondents were notified before the survey that all answers were recorded anonymously. The respondents were told at the beginning of the telephone survey that they could terminate the survey anytime if they want to. The respondents were also notified before the survey that they would have a chance to win a 10-dollar gift card if they fully participated.

@@

Three month after the implementation of our intervention, we conducted our formal survey using the same sampling frame and sampling methods. Please see the sampling details and the response rate in the table (Table \@ref(tab:presamplingtable)) below, our sampled population included 45 valid responses in Brantford and 48 collected from Peterborough, among which, 5 were collected via the online survey in Brantford and 7 were collected via the online survey in Peterborough.

```{r samplingtable, fig.cap='Sampling Results Summary', echo=FALSE }
tibble(
  'City' = list('Brantford', 'Peterborough'), 
  'Survey Administered' = list(100, 100),
  'Total Sample Collected' = list(45, 48),
  'Telephone Survey' = list(40, 41),
  'Online Survey' = list(5, 7)
) %>%
  kbl(caption = 'Sampling Results Summary',
      align = "cccc") %>%
  kable_styling(latex_options = "hold_position")
```

The two surveys incurred a total cost of \$10,000 CAD, which was mainly used to hire and train staff to conduct the telephone survey, collect data and administer the online survey if applicable. A small portion of the cost (\$500 CAD) was used for online survey technology support and database maintenance. We also gave out 50 gift cards with a total worth (\$500 CAD) as incentives for higher response rate.

## Main Survey Results
```{r, include=FALSE}
# read in survey data
simulated_dataset<-read_csv(here('outputs/survey/simulated_restaurant_data.csv'))
```

@@

The above figure (Figure \@ref(fig:question3)) shows that restaurants in Peterborough had lower rates of layoff compared to restaurants in Brantford, suggesting families in Peterborough to have experienced better financial situations and probably less stress.

@@

```
Although from Figure \@ref(fig:q8), we a pervasive gloom among the Brantford restaurants, Figure \@ref(fig:q10) illustrates that even if the pandemic ends immediately, many restaurants in Brantford, especially quick-service restaurants, have indicated that it is difficult to return to the state of operation before the COVID-19 outbreak. While for Peterborough, although some of them suspect the recovery will take more than a year to recoup with the previous situation, most of them indicated this period would be less than a year.

In addition, survey results about expected recovery time back to pre-COVID situation (Figure \@ref(fig:q10)) illustrate that even if the pandemic ends immediately, most restaurants in Brantford, especially quick-service restaurants, hold negative views to recover back to pre-COVID operating situation in a short period of time. For restaurants in Peterborough, despite some suspicions to recoup with the pre-COVID situation in more than a year, most of them expected to recovery in less than a year. Therefore, we believe that the continuing shutdown in Brantford have harmed expectations of future operation possibilities in Brantford, while the reopening in Peterborough has built confidence to resolve current operation difficulties.

```{r}
simulated_dataset %>%
  tabyl(q1_city, q11_rate_support)%>%
  adorn_totals("row")%>%
  adorn_totals("col")%>%
  adorn_percentages("row")%>%
  adorn_pct_formatting(digits=2)%>%
  adorn_ns()%>%
  adorn_title("combined",row_name = "City", col_name = "Rating")%>%
  kable("html", align = 'ccccccc', caption = 'Number or percentage of each rating level')%>%
  kable_styling(full_width = F)
```

 As an agent to gather relevant feedback to the Ontario Government, we asked restaurants how they found the governmental support programs so far. According to Table, both cities stated the effectiveness of the governmental support programs, nevertheless, this figure also implies to their satisfaction and also appreciation for these programs by business entities in both cities.
```{r}
# use
data.frame(table(simulated_dataset$q13_CEWS,simulated_dataset$q1_city))%>%
#  mutate(Var1 = fct_relevel(Var1 ,"Prefer not to say","More than $50,000","$40,000 ~ $50,000","$20,000 ~ $40,000","$0 ~ $20,000")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
    geom_bar(stat = "identity", width = 0.8, position = position_dodge())+
  ylab("Number of responses")+
  xlab("Options")+
  ggtitle("Restaurants receiving the Canada Emergency Wage Subsidy(CEWS)")+
  guides(fill=guide_legend(title="City"))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 0))+
  coord_flip()
```

@@

```{r}
data.frame(table(simulated_dataset$q14_loans,simulated_dataset$q1_city))%>%
  mutate(Var1 = fct_relevel(Var1 ,"Prefer not to say","More than $50,000","$40,000 ~ $50,000","$20,000 ~ $40,000","$0 ~ $20,000")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  geom_bar(stat = "identity", width = 0.8, position = position_dodge())+
  ylab("Number of responses")+
  xlab("Grant value")+
  ggtitle("How much loans have you applied for the COVID-19 related reasons?")+
  guides(fill=guide_legend(title="City"))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 0))+
  coord_flip()

@@

## Determine the Treatment Group and Control Group

@@

* The mindful comparison between cities' similarities on employment, population density, etc. ensures the internal validity of our experiment so that we could accurately estimate the true impact of shutdown restrictions on restaurant. However, our shutdown effect evaluation may not be generalized to the entire population of interest, i.e. all restaurants in Ontario such as township regions or cosmopolitan city Toronto due to different levels of competitions, operating costs, etc.

## Concerns about Sampling Method

@@

* Since the results of the presurvey suggest apparent similarities between the restaurants in Peterborough and those in Brantford, we assume the surveyed restaurants to have similar **qualities**. However, the attributes of the restaurants (size, employments, service type, etc) were not controlled in our analysis. In other words, we believe the response rate within the different restaurant subgroups to be the same. Variations in the actual response between restaurants in two cities are unknown and may impact the internal validity of the results.

@@

## Survey Results and Policy Implication
-	Staffing implications 
*	Policy implications: The Government of Ontario should continue the existing supports provided to restaurants during the lockdown and consider additional support measures to mitigate the risk of increased unemployment. Reopening the restaurants may lead to increased employment in the corresponding area. 
-	Restaurant sales implementation 
*	The Government should expect better sales for reopened restaurants. However, it is important to acknowledge that economy would take a long time to reach pre-pandemic performance. Continued support to restaurants may still be needed. 
-	Future projects of restaurants  
*	Rebuilding business confidence is a crucial aspect of short-term business recovery and long-term economic stability. The Government should carefully monitor business projection and equip businesses with tailored supports based on their recovery needs. 
-	Striking a balance between COVID-19 cases

```{r}
# readin the covid tracking data
covid_tracking <- readr::read_csv(here::here("inputs/data/simulated_covid_data.csv"))
```

>>>>

header-includes:
 - \usepackage{float}
 - \floatplacement{figure}{H}
 - \floatplacement{table}{H}

@@

```{r setup, include=FALSE}
# May install the following displaying tables
# Reference: https://bookdown.org/yihui/bookdown/html-widgets.html
# install.packages("webshot")
# webshot::install_phantomjs()

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(bookdown)
library(janitor)
library(stringr)
library(ggrepel)
library(here)
library(janitor)
library(kableExtra)
library(finalfit)
```

@@

&nbsp; &nbsp; Our data explores shutdown effects on restaurant businesses in Ontario. We analyzed it using R [@citeR], and packages `tidyverse` [@cite_tidyverse], `stringr`[@cite_stringr],  `here` [@cite_here], `ggpubr`[@cite_ggpubr], `janitor`[@cite_janitor] . We used R packages `cansim` [@cite_cansim], `httr`[@cite_httr], `xml2`[@cite_xml2] to decide our intervention and sampling method, and packages `bookdown` [@cite_bookdown], `kableExtra`[@cite_kableextra], `ggrepel`[@cite_ggrepel], `finalfit`[@cite_finalfit] to format the document. We referenced `Impact Evaluation in Practice`[@cite_impact] and `Sampling Theory and Practice`[@cite_sampletheory] to design our experiment.

@@

&nbsp; &nbsp; In September, 2020, we at Petit Poll conducted the first survey on restaurants in Peterborough and Brantford to examine their business conditions before the experiment. Starting from October, 2020, the Ontario government reopened Peterborough while keeping shutdown restrictions on the rest of Ontario regions. After a three month period in January 2021, we at Petit Poll started to conduct survey on restaurants in Peterborough and Brantford about their business performance from October to December. Survey results were released in February 2021.

@@

&nbsp; &nbsp; As we would like to understand the effect of COVID shutdowns on restaurant businesses in Ontario, we considered all restaurants in Ontario as our target population. According to our intervention, the frame population are all restaurants in Peterborough and Brantford. And approaches to obtain the sampled population is described as follow:

1. Scrape postal codes of all restaurants located within the circle of the Brantford City Centre to a 40 km radius on Yelp. And then apply the same method for Peterborough.^[Land areas of both Brantford and Peterborough are less than 100 kilometer squares, so searching restaurants 40 km within the city center almost ensures to include all restaurants in Peterborough and Brantford]

@@

3. Compare these two lists and eliminate restaurants with postal codes outside of Peterborough CMA and Brantford CMA to get the list of all restaurants in Brantford and Peterborough as the frame population.

4. Randomly select 100 restaurants from Brantford and 100 restaurants from Peterborough to form the sampled population.

&nbsp; &nbsp; The first survey was conducted before the decision on intervention to ensure our assumptions about restaurant similarities between Peterborough and Brantford are based on concert evidence. We asked selected respondents about the restaurants' employment, average monthly sales, nature and category of restaurant, restaurant capacity to examine restaurant similarities between these two cities.

```{r presamplingtable, fig.cap='Sampling Summary of the First Survey', echo=FALSE }
tibble(
  'City' = list('Brantford', 'Peterborough'), 
  'Survey Administered' = list(100, 100),
  'Total Sample Collected' = list(50, 55),
  'Telephone Survey' = list(43, 47),
  'Online Survey' = list(7, 8)
) %>%
  kbl(caption = 'Sampling Summary of the First Survey',
      align = "cccc") %>%
  kable_styling(latex_options = "hold_position")
```

&nbsp; &nbsp; We used telephone interviews and online questionnaire methods to administer the survey. The telephone interview served as the primary channel and an online survey option was provided when respondents could not or prefer not to be interviewed over the phone. As shown in the above table (Table \@ref(tab:presamplingtable)), our sampled population included 50 valid responses in Brantford and 55 collected from Peterborough, among which, 7 were collected via the online survey in Brantford and 8 were collected via the online survey in Peterborough. Surveys that were not answered or have incomplete answers were treated as nonresponses and were not included in the dataset for further analysis. The surveys were administered in following steps:

1. Scraped phone numbers of restaurants from Yelp while we determined the frame population.

@@

3. When the call was connected, the survey staff would provide a brief introduction of the survey purpose and indicate that the survey was sponsored by the Ontario government.^[Scripts used to approach respondents are included in Appendix A and B ].

@@

6. To address privacy concerns and to increase the response rate, the respondents were notified before the survey that all answers were recorded anonymously. The respondents were told at the beginning of the telephone survey that they could terminate the survey anytime if they want to. The respondents were also notified before the survey that they would have a chance to win a 50-dollar gift card if they fully participated.
```{r survey1data, include=FALSE}
simulated_survey1_dataset <- read.csv( here('outputs/survey/simulated_first_survey.csv'))
```

```{r survey1, tab.cap="Summary Statistics for the First Survey", echo = FALSE, message = FALSE, warning=FALSE}
f_survey <- simulated_survey1_dataset %>% 
  summary_factorlist("q1_city", c("q3_category","q5_nature","q2_employees","q4_sales","q6_customers"),
                     p = FALSE, 
                     add_dependent_label = TRUE,
                     dependent_label_prefix = "",
                     add_col_totals = TRUE,
                     include_row_missing_col = FALSE,
                     col_totals_rowname = "",
                     total_col = FALSE) %>%
  rename(
    Responses = ` `
  )
f_survey$q1_city <- NULL
#  f_survey$Responses[c(1:5,9,8,6,7,10,11,13,14,15,12,16,21,17,18,19,20,22),] # Manually change the order of the factor levels to have a better visualization 
 f_survey <- f_survey[c(1:5,9,8,6,7,10,11,13,14,15,12,16,21,17,18,19,20,22),] %>% 
    mutate(` ` = c("Total Samples","Restaurant Category","","Restaurant Nature","","Number of Employees","","","","","Average Monthly Sales","","","","","","Restaurant Capacity","","","","",""))

kable( f_survey[,c(4,1,2,3)],caption = "Summary Statistics of the First Survey",
       booktabs = TRUE,
       valign='t') 
```

@@

After three months of our intervention, we conducted our formal survey using the same sampling frame and sampling methods. Please see the sampling details and the response rate in the table (Table \@ref(tab:presamplingtable)) below, our sampled population included 45 valid responses in Brantford and 48 collected from Peterborough, among which, 5 were collected via the online survey in Brantford and 7 were collected via the online survey in Peterborough.

```{r samplingtable, fig.cap='Sampling Results Summary', echo=FALSE }
tibble(
  'City' = list('Brantford', 'Peterborough'), 
  'Survey Administered' = list(100, 100),
  'Total Sample Collected' = list(45, 48),
  'Telephone Survey' = list(40, 41),
  'Online Survey' = list(5, 7)
) %>%
  kbl(caption = 'Sampling Summary of the Second Summary',
      align = "cccc") %>%
  kable_styling(latex_options = "hold_position")
```

The two surveys incurred a total cost of \$15,000 CAD, which was mainly used to hire and train staff to conduct the telephone survey, collect data and administer the online survey if applicable. A small portion of the cost (\$500 CAD) was used for online survey technology support and database maintenance. We also gave out 50 gift cards with a total worth (\$500 CAD) as incentives for higher response rate.

## Main Survey Results
```{r, include=FALSE}
# read in survey data
simulated_dataset<-read_csv(here('outputs/survey/simulated_second_survey.csv'))
```

@@

Compared to Peterborough, restaurants in Brantford have much higher number of laid-offs (Figure \@ref(fig:question3)), suggesting the continuing shutdown greatly affected the 
The above figure (Figure \@ref(fig:question3)) shows that restaurants in Peterborough had lower rates of layoff compared to restaurants in Brantford, suggesting families in Peterborough to have experienced better financial situations and probably less stress.

@@

In addition, survey results about expected recovery time back to pre-COVID situation (Figure \@ref(fig:q10)) illustrate that even if the pandemic ends immediately, most restaurants in Brantford, especially quick-service restaurants, hold negative views to recover back to pre-COVID operating situation in a short period of time. For restaurants in Peterborough, despite some suspicions to recoup with the pre-COVID situation in more than a year, most of them expected to recovery in less than a year. Therefore, we believe that the continuing shutdown in Brantford have harmed expectations of future operation possibilities in Brantford, while the reopening in Peterborough could build confidence to alleviate current operation difficulties.

```{r q11, tab.cap="Satisfaction Level of Government Financial Support Programs on Restaurants", echo = FALSE, message = FALSE, warning=FALSE}
simulated_dataset %>%
  tabyl(q1_city, q11_rate_support)%>%
  adorn_totals("row")%>%
  adorn_totals("col")%>%
  adorn_percentages("row")%>%
  adorn_pct_formatting(digits=2)%>%
  adorn_ns()%>%
  adorn_title("combined",row_name = "City", col_name = "Rating")%>%
  rename(
    `Terrible` = `1`,
    `Poor` = `2`,
    `Fair` = `3`,
    `Good` = `4`,
    `Excellent` = `5`
  ) %>% 
  kable("html", align = 'ccccccc', caption = 'Satisfaction Level of Government Financial Support Programs on Restaurants') %>% 
  kable_styling(full_width = F)
```

 As agents to gather relevant feedbacks to the Ontario Government, we asked restaurants how they found the governmental support programs so far. The survey results (Table \@ref(tab:q11)) shows that both cities generally stated the effectiveness of the governmental support programs, **nevertheless, this figure  also implies to their satisfaction and also appreciation for these programs by business entities in both cities.**

```{r q14, fig.align='center', fig.cap="Restaurants receiving the Canada Emergency Wage Subsidy(CEWS)", fig.width=8, fig.height=4, echo= FALSE, message=FALSE, warning=FALSE}
# use
data.frame(table(simulated_dataset$q13_CEWS,simulated_dataset$q1_city))%>%
#  mutate(Var1 = fct_relevel(Var1 ,"Prefer not to say","More than $50,000","$40,000 ~ $50,000","$20,000 ~ $40,000","$0 ~ $20,000")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
    geom_bar(stat = "identity", width = 0.8, position = position_dodge())+
  ylab("Number of responses")+
  xlab("Options")+
  ggtitle("Restaurants receiving the Canada Emergency Wage Subsidy(CEWS)")+
  guides(fill=guide_legend(title="City"))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 0))+
  coord_flip()
```

@@

```{r q15, fig.align='center', fig.cap="Loan amount applied", fig.width=8, fig.height=4, echo= FALSE, message=FALSE, warning=FALSE}
data.frame(table(simulated_dataset$q14_loans,simulated_dataset$q1_city))%>%
  mutate(Var1 = fct_relevel(Var1 ,"Prefer not to say","More than $50,000","$40,000 ~ $50,000","$20,000 ~ $40,000","$0 ~ $20,000")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  geom_bar(stat = "identity", width = 0.8, position = position_dodge())+
  ylab("Number of responses")+
  xlab("Grant value")+
  ggtitle("How much loans have you applied for the COVID-19 related reasons?")+
  guides(fill=guide_legend(title="City"))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 0))+
  coord_flip()

@@

## Survey Results and Policy Implication
Staffing implications 
* Our survey found restaurants in Peterborough had lower rates of layoff compared to restaurants in Brantford, suggesting that reopening restaurants may lead to more employment in the industry. 
*	Policy implications: The Government of Ontario should continue the existing supports provided to restaurants during the lockdown and consider additional support measures to mitigate the risk of increased unemployment. Reopening the restaurants may lead to increased employment in the corresponding area. 
Restaurant sales implementation 
* We found a considerable increase in sales in Peterborough during the reopening period and a slight decrease in Brantford, possibly due to seasonal fluctuations. The results showed that reopening could lead to more business revenue for the restaurants. 
*	Policy implication: The Government should expect better sales for reopened restaurants. However, it is important to acknowledge that the economy would take a long time to reach pre-pandemic performance. Continued support to restaurants may still be needed in the near future.  
Future projects of restaurants  
* Restaurants in Peterborough showed more optimism about their future success, while restaurants in Brantford were less confident. The reopening and the subsequent increase in revenue had instilled hope for the businesses.  
* Government implication: Rebuilding business confidence is a crucial aspect of short-term business recovery and long-term economic stability. The Government should carefully monitor business projection and equip businesses with tailored supports based on their recovery needs. 
Dependence on government support 
* Compared with Peterborough, Bradford restaurants showed higher dependence on government programs, possibly due to the lack of strong revenue streams. We also found that restaurants in Peterborough are leveraging the CEWS and government loan programs more, which may be a result of increased employment and confidence in their abililties to pay back the loans.  
* Policy implication: When implementing reopening policies, the government could leverage more direct non-repayable funding support and reasonably reduce the use of indirect low-interest loans for the reopened business. 
Striking a balance between COVID-19 cases

```{r coviddata, include=FALSE}
# readin the covid tracking data
covid_tracking <- readr::read_csv(here::here("inputs/data/simulated_covid_data.csv"))
```

## Limitations in Determine the Treatment Group and Control Group

@@

* The mindful comparison between cities' similarities on employment, population density, etc. ensures the internal validity of our experiment so that we could accurately estimate the true impact of shutdown restrictions on restaurant. However, our shutdown effect evaluations may not be generalized to the entire population of interest, i.e. all restaurants in Ontario such as township regions or cosmopolitan city Toronto due to different levels of competitions, operating costs, etc.

## Limitations in Sampling Method

@@

* Since results of the first survey before implementing our experiment suggest apparent similarities between the restaurants in Peterborough and those in Brantford, we assume the surveyed restaurants to have similar **qualities**. However, the attributes of the restaurants (size, employments, service type, etc) were not controlled in our analysis. In other words, we believe the response rate within the different restaurant subgroups to be the same. Variations in the actual response between restaurants in two cities are unknown and may impact the internal validity of the results.

@@

**High response rate:**
